# Data transformation

```{r setup, include=FALSE}
# this prevents package loading message from appearing in the rendered version
knitr::opts_chunk$set(warning = FALSE, message = FALSE,
                      echo = TRUE)

library(tidyverse)
library(patchwork)
library(readxl)
library(zoo)
library(lubridate)
```

For crime data of all types from 2017 to 2021, we performed the following steps to clean the data:

1. Select out and rename the columns of data we are interested in: Precinct, Date, Type of Crime, and Subtype of Crime. 
2. Find out corresponding Borough based on Precinct
3. The final output csv consists of the above 5 features while each row represents a crime reported.

The following code demonstrates the workflow we applied to crime data in 2021. We also applied the same workflow to the data in 2017 - 2020 separately in order to save space.

```{r, crime data 2017-2021, eval=FALSE}

df <- read_csv("NYPD_Complaint_Data_Current__Year_To_Date_.csv", show_col_types = FALSE)

old_names <- c('ADDR_PCT_CD', 'CMPLNT_FR_DT', 'LAW_CAT_CD', 'OFNS_DESC')
new_names <- c('pct', 'date', 'level', 'type')
df <- df[,old_names]
names(df) <- new_names

df$month <- unlist(lapply(df$date, function(x) as.numeric(str_split(x, '/')[[1]][1])))
df$year <- unlist(lapply(df$date, function(x) as.numeric(str_split(x, '/')[[1]][3])))
df <- df[df$year == 2021,]

find_boro <- function(pct){
  if (pct %in% c(40,41,42,43,44,45,46,47,48,49,50,52)) {
    return('BRONX')
  } else if (pct %in% c(60,61,62,63,66,67,68,69,70,71,72,73,75,76,77,78,79,81,83,84,88,90,94)) {
    return('BROOKLYN')
  } else if (pct %in% c(1,5,6,7,9,10,13,14,17,18,19,20,22,23,24,25,26,28,30,32,33,34)) {
    return('MANHATTAN')
  } else if (pct %in% c(100,101,102,103,104,105,106,107,108,109,110,111,112,113,114,115)) {
    return('QUEENS')
  } else if (pct %in% c(120,121,122,123)) {
    return('STATEN ISLAND')
  } 
}

df$borough <- unlist(lapply(df$pct, function(x) find_boro(x)))
write.csv(df,"crime_21.csv")

```

For park crime data from 2017 to 2021, we performed the following steps to clean the data:

1. Use the same naming of columns for all files
2. Fix data types 
3. Aggregate rows based on boroughs and compute the number of parks for each borough
4. Specify quarter and year
5. Combine park crime in all quarters from 2017 to 2021 into one single csv

```{r, read in park crime data 2017 - 2021, eval=FALSE, include=FALSE}
q1_17 <- read_excel('nyc-park-crime-stats-q1-2017.xlsx', skip = 3)
q2_17 <- read_excel('nyc-park-crime-stats-q2-2017.xlsx', skip = 3)
q3_17 <- read_excel('nyc-park-crime-stats-q3-2017.xlsx', skip = 3)
q4_17 <- read_excel('nyc-park-crime-stats-q4-2017.xlsx', skip = 3)

q1_18 <- read_excel('nyc-park-crime-stats-q1-2018.xlsx', skip = 3)
q2_18 <- read_excel('nyc-park-crime-stats-q2-2018.xlsx', skip = 4)
q3_18 <- read_excel('nyc-park-crime-stats-q3-2018.xlsx', skip = 3)
q4_18 <- read_excel('nyc-park-crime-stats-q4-2018.xlsx', skip = 3)

q1_19 <- read_excel('nyc-park-crime-stats-q1-2019.xlsx', skip = 3)
q2_19 <- read_excel('nyc-park-crime-stats-q2-2019.xlsx', skip = 3)
q3_19 <- read_excel('nyc-park-crime-stats-q3-2019.xlsx', skip = 3)
q4_19 <- read_excel('nyc-park-crime-stats-q4-2019.xlsx', skip = 3)

q1_20 <- read_excel('nyc-park-crime-stats-q1-2020a.xlsx', skip = 3)
q2_20 <- read_excel('nyc-park-crime-stats-q2-2020.xlsx', skip = 3)
q3_20 <- read_excel('nyc-park-crime-stats-q3-2020.xlsx', skip = 3)
q4_20 <- read_excel('nyc-park-crime-stats-q4-2020.xlsx', skip = 3)

q1_21 <- read_excel('nyc-park-crime-stats-q1-2021.xlsx', skip = 4)
q2_21 <- read_excel('nyc-park-crime-stats-q2-2021.xlsx', skip = 3)
q3_21 <- read_excel('nyc-park-crime-stats-q3-2021.xlsx', skip = 3)
```

```{r, park crime data 2017 - 2021, eval=FALSE}

agg <- function(df, q, y){
  names(df) <- c('PARK', 'BOROUGH', 'SIZE (ACRES)', 'CATEGORY', 'MURDER', 'RAPE', 
                 'ROBBERY', 'FELONY ASSAULT', 'BURGLARY', 'GRAND LARCENY',
                 'GRAND LARCENY OF MOTOR VEHICLE', 'TOTAL')
  
  df <- df[1:(nrow(df)-2),]
  
  features <- c('MURDER', 'RAPE', 'ROBBERY', 'FELONY ASSAULT', 'BURGLARY', 'GRAND LARCENY',
                'GRAND LARCENY OF MOTOR VEHICLE', 'TOTAL')
  
  df[,features] <- lapply(df[,features], function(x) as.numeric(x))
  
  df <- df %>% mutate(num_parks = 1)
  
  df <- as_tibble(aggregate(x = df[,c(features,c('num_parks'))], by = list(df$BOROUGH), FUN = sum))
  
  
  names(df)[1] <- 'BOROUGH'
  
  df$QUARTER <- q
  
  df$YEAR <- y
  
  return(df)
}

# q1_17 represents the data in the first quarter of 2017

df <- agg(q1_17, 1, 2017)
df <- rbind(df, agg(q2_17, 2, 2017))
df <- rbind(df, agg(q3_17, 3, 2017))
df <- rbind(df, agg(q4_17, 4, 2017))
df <- rbind(df, agg(q1_18, 1, 2018))
df <- rbind(df, agg(q2_18, 2, 2018))
df <- rbind(df, agg(q3_18, 3, 2018))
df <- rbind(df, agg(q4_18, 4, 2018))
df <- rbind(df, agg(q1_19, 1, 2019))
df <- rbind(df, agg(q2_19, 2, 2019))
df <- rbind(df, agg(q3_19, 3, 2019))
df <- rbind(df, agg(q4_19, 4, 2019))
df <- rbind(df, agg(q1_20, 1, 2020))
df <- rbind(df, agg(q2_20, 2, 2020))
df <- rbind(df, agg(q3_20, 3, 2020))
df <- rbind(df, agg(q4_20, 4, 2020))
df <- rbind(df, agg(q1_21, 1, 2021))
df <- rbind(df, agg(q2_21, 2, 2021))
df <- rbind(df, agg(q3_21, 3, 2021))

write.csv(df,"park_crime.csv")
```