# Missing values

Here is the function for plotting missing values

```{r, pacakges missing values, include=FALSE}
library(tidyverse)
library(patchwork)
```

```{r, plot function}
plot_missing <- function(data, percent = FALSE) {
  if (percent == FALSE) {
    cols <- colSums(is.na(data)) %>% sort(decreasing = TRUE)
    missing_label = 'missing'
  } else {
    cols <- colSums(is.na(data))
    if (sum(cols) > 0) {
      cols <- cols/sum(cols)*100
    }
    cols <- cols %>% sort(decreasing = TRUE)
    missing_label = '%missing'
  }
  
  pu <- ggplot() +
    geom_bar(aes(x = names(cols), y = cols), fill = 'cornflowerblue', stat = 'identity') +
    ggtitle('Missing value patterns') +
    scale_x_discrete(limits = names(cols)) +
    ylab(missing_label) +
    theme_bw() +
    theme(axis.title.x = element_blank(), panel.grid.major.x = element_blank(), axis.text.x = element_text(angle = 90))
  
  missing_patterns <- data.frame(is.na(data[, names(cols)])) %>%
    group_by_all() %>%
    count(name = 'count', sort = TRUE) %>%
    ungroup()
  
  row_count <- missing_patterns$count
  missing_patterns <- missing_patterns[,names(missing_patterns) != 'count'] * 1
  
  complete_rows <- apply(missing_patterns, 1, function(row) all(row == 0 ))
  missing_patterns[complete_rows,] <- 2
  
  df <- missing_patterns %>% 
    mutate(id = row_number()) %>%
    gather(key, value, -id)
  
  pm <- ggplot(df, aes(key, fct_rev(as.factor(id)), fill = as.factor(value))) +
    geom_tile(color = 'white') +
    scale_fill_manual(values = c('0' = 'gray', '1' = 'blueviolet', '2' = 'black')) +
    scale_x_discrete(limits = names(cols)) + 
    ylab('missing pattern') +
    xlab('variable') +
    theme_bw() +
    theme(legend.position = 'none', axis.text.x = element_text(angle = 90))
  
  if (percent == TRUE) {
    row_count <- row_count/sum(row_count) * 100
  } 
  
  r <- data.frame('row count' = row_count, 'r_id' = 1:length(row_count), 'complete' = complete_rows*1)
  pr <- ggplot(r, aes(fct_rev(as.factor(r_id)), row_count, fill = as.factor(complete))) +
    geom_bar(stat = 'identity') +
    scale_fill_manual(values = c('0' = 'cornflowerblue', '1' = 'blue')) +
    coord_flip() + 
    theme_bw() +
    theme(axis.title.y = element_blank(), panel.grid.major.y = element_blank(), legend.position = 'none')
  
  layout <- '
  AAA#
  BBBC
  BBBC
  BBBC
  '
  
  pu + pm + pr + plot_layout(design = layout)
}
```

Both the raw data files and the transformed data files do not contain missing values. The first reason is that these data are safety related, and thus must be well documented in detail. The other reason is that we are focusing on relatively common features/columns, such as precinct, borough,crime type, and crime subtype. Some columns in raw data do contain missing values (Ex. latitudes and longitudes), but they are of little interest. Considering that the size of raw data is too large to present, we only use one of our cleaned data sets, park crime data, for demonstration here.

The first plot shows "count of missing values" while the second plot shows "percentage of missing values".

**Park Crime Data Missing Values**
```{r, read in data, include=FALSE}
url <- "https://raw.githubusercontent.com/KecheZhu/crimedata/main/data/park_crime.csv"
df <- read.csv(url)
df <- df[,names(df) != 'X']
names(df) <- c("Borough","Murder","Rape","Robbery","Felony_assault","Burglary","GL","GL_MV",
               "Total","Num_Parks","Quarter","Year")
```
```{r, plot1}
plot_missing(df, percent = FALSE)
```
```{r, plot2}
plot_missing(df, percent = TRUE)
```

